### 2022-06-02 chapters/fairness_background.tex
minor edits in many parts of the thesis

\part{Fairness in Machine Learning}
\chapter{Preliminaries}
We represent sets/vectors by bold letters, and the corresponding distributions by calligraphic letters. We express random variables in uppercase, and an assignment of a random variable in lowercase.

@@

\begin{example}
	Following~\cite{bastani2019probabilistic}, we consider a classifier that decides the hiring of employees based on three features: gender (sensitive), years of experience (non-sensitive), and college-participation (mediator). It is practical to consider that gender $ \in $ \{male, female\} can affect the college-participation of individuals, and all three features are determining factors for the hiring process. Let `male' be the most favored group by the classifier, for instance. Path-specific causal fairness ensures that a female candidate should be given a job offer with similar probability as a male candidate (by constraining $ \epsilon \approx 0 $). She,  however,  went to (participated in) college as if she were a male candidate while other non-mediator features such as  `years of experience' are the same.  Therefore, path-specific causal fairness measures the effect of gender on job offer, but ignores the effect of gender on whether candidates went to college.
\end{example}

@@

\subsection*{Stochastic Boolean Satisfiability (SSAT)}\label{fairness_justicia_sec:ssat}
Let $\mathbf{B}  = \{\bool_1, \dots, \bool_m\}  $ be a set of Boolean variables. A \textit{literal} is a variable $ \bool_i $ or its complement $ \neg \bool_i $. 
A propositional formula $\phi$ defined over $\mathbf{B}$ is in \textit{Conjunctive Normal Form (CNF)} if $\phi$   is  a conjunction of clauses and each clause is a disjunction of literals. 
%\red{DNF (disjunctive normal form) is the  complement of CNF where  the formula is a disjunction of clauses and each clause is  a conjunction of literals.} 
Let $ \sigma $ be a assignment to the  variables $ \bool_i \in \mathbf{B} $  such that $ \sigma(\bool_i) \in \{1, 0\} $. The propositional  \textit{satisfiability} problem (SAT) finds a satisfying assignment $ \sigma^* $ to all $ \bool_i \in  \mathbf{B} $ such that the formula $ \phi $ is evaluated to be $1$ (equivalently, true). 
In contrast to the SAT problem, the \textit{Stochastic Boolean Satisfiability} (SSAT) problem~\cite{littman2001stochastic} is computes the  probability of the satisfaction of the formula $\phi$ defined on \textit{quantified} Boolean variables. 
An SSAT formula is of the form
\begin{equation}\label{fairness_justicia_eq:ssat}
\Phi = Q_1\bool_1, \dots, Q_m \bool_m,\; \phi, 
\end{equation}
where $ Q_i \in \{\exists, \forall, \R^{p_i}\} $ is either of the existential ($\exists$), universal ($\forall$), or randomized ($\R^{p_i}$) quantifiers on $\bool_i$ and $\phi$ is a quantifier-free CNF formula. In case of a randomized quantifier $ \R^{p_i} $, $ p_i \in [0,1] $ is the probability of $ \bool_i $ being assigned to $ 1 $. In the SSAT formula $ \Phi $, the quantifier part $ Q_1\bool_1, \dots, Q_m \bool_m $ is known as the \textit{prefix} of the formula $ \phi $.  Let $ \bool $ be the outermost variable in the prefix. The semantics of SSAT formulas are defined recursively in the following.
\begin{enumerate}
	\item $ \Pr[\text{true}] = 1 $,  $ \Pr[\text{false}] = 0 $, 
	\item $ \Pr [\Phi] = \max_{\bool} \{\Pr[\Phi|_{\bool}], \Pr[\Phi|_{\neg \bool}]\}$ if $ \bool $ is existentially quantified ($ \exists $), 
	\item $ \Pr [\Phi] = \min_{\bool} \{\Pr[\Phi|_{\bool}], \Pr[\Phi|_{\neg \bool}]\} $ if $ \bool $ is universally quantified ($ \forall $), 
	\item $ \Pr [\Phi] = p\Pr[\Phi|_{\bool}] + (1-p) \Pr[\Phi|_{\neg \bool}] $ if $ \bool $ is randomized quantified ($\R^{p}$) with probability $p$ of being $\text{true}$,
\end{enumerate}
where $ \Phi|_{\bool} $ and $ \Phi|_{\neg \bool} $ denote the SSAT formulas derived by eliminating the outermost quantifier of $ \bool $  by substituting the value of $ \bool $ in the CNF $ \phi $ with $ 1 $ and $ 0 $, respectively. In this paper, we focus on two specific types of SSAT formulas:  \textit{random-exist} (RE) SSAT and \textit{exist-random} (ER) SSAT. In the ER-SSAT (resp.\ RE-SSAT) formula, all existentially (resp.\ randomized) quantified variables are followed by randomized (resp.\ existentially) quantified variables in the prefix.

@@

\subsection*{Bayesian Network}
In general, a Probabilistic Graphical Model~\cite{koller2009probabilistic}, and specifically a \textit{Bayesian network}~\cite{pearl1985bayesian,chavira2008probabilistic}, encodes the dependencies and conditional independence between a set of random variables. In this paper, we leverage an access to a Bayesian network on $ \nonsensitive \cup \sensitive $ that represents the joint distribution on them. 	A Bayesian network is denoted by a pair $ (\graph, \theta)$, where $ \graph \triangleq (\mathbf{V}, \mathbf{E}) $ is a DAG (Directed Acyclic Graph), and $\theta$ is a set of parameters encoding the conditional probabilities induced by the joint distribution under investigation. Each vertex $V_i \in \mathbf{V}$ corresponds to a random variable. Edges $ \mathbf{E} \in \mathbf{V} \times \mathbf{V} $ imply conditional dependencies among variables. For each variable $ V_i \in \mathbf{V} $, let $ \parent(V_i) \subseteq \mathbf{V} \setminus \{V_i\} $ denote the set of parents of $ V_i $. Given $\parent(V_i)$ and parameters $\theta$, $ V_i $ is independent of its other non-descendant variables in $\graph$. Thus, for the assignment $ v_i $ of $ V_i $ and $ \mathbf{u} $ of $ \parent(V_i) $, the aforementioned semantics of a Bayesian network encodes the joint distribution of $\mathbf{V}$ as:

>>>>

\part{Fairness in Machine Learning}
\chapter{Preliminaries}
In this chapter, we discuss preliminaries in fairness in machine learning, followed by methods to verify fairness and identify the sources of unfairness. We represent sets/vectors by bold letters, and the corresponding distributions by calligraphic letters. We express random variables in uppercase, and an assignment of a random variable in lowercase.

@@

\begin{example}
	\normalfont
	Following~\cite{bastani2019probabilistic}, we consider a classifier that decides the hiring of employees based on three features: gender (sensitive), years of experience (non-sensitive), and college-participation (mediator). It is practical to consider that gender $ \in $ \{male, female\} can affect the college-participation of individuals, and all three features are determining factors for the hiring process. Let `male' be the most favored group by the classifier, for instance. Path-specific causal fairness ensures that a female candidate should be given a job offer with similar probability as a male candidate (by constraining $ \epsilon \approx 0 $). She,  however,  went to (participated in) college as if she were a male candidate while other non-mediator features such as  `years of experience' are the same.  Therefore, path-specific causal fairness measures the effect of gender on job offer, but ignores the effect of gender on whether candidates went to college.
\end{example}

@@

\subsection*{Stochastic Boolean Satisfiability (SSAT)}\label{fairness_justicia_sec:ssat}
Let $\mathbf{B}  = \{\bool_1, \dots, \bool_m\}  $ be a set of Boolean variables. A \textit{literal} is a variable $ \bool_i $ or its complement $ \neg \bool_i $. 
A propositional formula $\phi$ defined over $\mathbf{B}$ is in \textit{Conjunctive Normal Form (CNF)} if $\phi$   is  a conjunction of clauses and each clause is a disjunction of literals. 
%\red{DNF (disjunctive normal form) is the  complement of CNF where  the formula is a disjunction of clauses and each clause is  a conjunction of literals.} 
Let $ \sigma $ be a assignment to the  variables $ \bool_i \in \mathbf{B} $  such that $ \sigma(\bool_i) \in \{1, 0\} $. The propositional  \textit{satisfiability} problem (SAT) finds a satisfying assignment $ \sigma^* $ to all $ \bool_i \in  \mathbf{B} $ such that the formula $ \phi $ is evaluated to be $1$ (equivalently, true). 
In contrast to the SAT problem, the \textit{Stochastic Boolean Satisfiability} (SSAT) problem~\cite{littman2001stochastic} is computes the  probability of the satisfaction of the formula $\phi$ defined on \textit{quantified} Boolean variables. 
An SSAT formula is of the form
\begin{equation}\label{fairness_justicia_eq:ssat}
\Phi = Q_1\bool_1, \dots, Q_m \bool_m,\; \phi, 
\end{equation}
where $ Q_i \in \{\exists, \forall, \R^{p_i}\} $ is either of the existential ($\exists$), universal ($\forall$), or randomized ($\R^{p_i}$) quantifiers on $\bool_i$ and $\phi$ is a quantifier-free CNF formula. In case of a randomized quantifier $ \R^{p_i} $, $ p_i \in [0,1] $ is the probability of $ \bool_i $ being assigned to $ 1 $. In the SSAT formula $ \Phi $, the quantifier part $ Q_1\bool_1, \dots, Q_m \bool_m $ is known as the \textit{prefix} of the formula $ \phi $.  Let $ \bool $ be the outermost variable in the prefix. The semantics of SSAT formulas are defined recursively in the following.
\begin{enumerate}
	\item $ \Pr[\text{true}] = 1 $,  $ \Pr[\text{false}] = 0 $, 
	\item $ \Pr [\Phi] = \max_{\bool} \{\Pr[\Phi|_{\bool}], \Pr[\Phi|_{\neg \bool}]\}$ if $ \bool $ is existentially quantified ($ \exists $), 
	\item $ \Pr [\Phi] = \min_{\bool} \{\Pr[\Phi|_{\bool}], \Pr[\Phi|_{\neg \bool}]\} $ if $ \bool $ is universally quantified ($ \forall $), 
	\item $ \Pr [\Phi] = p\Pr[\Phi|_{\bool}] + (1-p) \Pr[\Phi|_{\neg \bool}] $ if $ \bool $ is randomized quantified ($\R^{p}$) with probability $p$ of being $\text{true}$,
\end{enumerate}
where $ \Phi|_{\bool} $ and $ \Phi|_{\neg \bool} $ denote the SSAT formulas derived by eliminating the outermost quantifier of $ \bool $  by substituting the value of $ \bool $ in the CNF $ \phi $ with $ 1 $ and $ 0 $, respectively. In this thesis, we focus on two specific types of SSAT formulas:  \textit{random-exist} (RE) SSAT and \textit{exist-random} (ER) SSAT. In the ER-SSAT (resp.\ RE-SSAT) formula, all existentially (resp.\ randomized) quantified variables are followed by randomized (resp.\ existentially) quantified variables in the prefix.

@@

\subsection*{Bayesian Network}
In general, a Probabilistic Graphical Model~\cite{koller2009probabilistic}, and specifically a \textit{Bayesian network}~\cite{pearl1985bayesian,chavira2008probabilistic}, encodes the dependencies and conditional independence between a set of random variables. In this chapter, we leverage an access to a Bayesian network on $ \nonsensitive \cup \sensitive $ that represents the joint distribution on them. 	A Bayesian network is denoted by a pair $ (\graph, \theta)$, where $ \graph \triangleq (\mathbf{V}, \mathbf{E}) $ is a DAG (Directed Acyclic Graph), and $\theta$ is a set of parameters encoding the conditional probabilities induced by the joint distribution under investigation. Each vertex $V_i \in \mathbf{V}$ corresponds to a random variable. Edges $ \mathbf{E} \in \mathbf{V} \times \mathbf{V} $ imply conditional dependencies among variables. For each variable $ V_i \in \mathbf{V} $, let $ \parent(V_i) \subseteq \mathbf{V} \setminus \{V_i\} $ denote the set of parents of $ V_i $. Given $\parent(V_i)$ and parameters $\theta$, $ V_i $ is independent of its other non-descendant variables in $\graph$. Thus, for the assignment $ v_i $ of $ V_i $ and $ \mathbf{u} $ of $ \parent(V_i) $, the aforementioned semantics of a Bayesian network encodes the joint distribution of $\mathbf{V}$ as:

### 2022-06-02 chapters/interpretability/imli/baseline.tex
minor edits in many parts of the thesis

\subsection{Description of Variables}
\label{interpretability_imli_sec:variables} 
We initially preprocess $ \mathbf{X} $  to account for the negation of  Boolean features while learning a classifier.\footnote{This preprocessing is similarly applied in  \cite{malioutov2013exact}.} In the preprocessing step, we negate each column in $ \mathbf{X} $ to a new column and append it to $ \mathbf{X} $. For example, if ``$ \text{age }\ge 25 $'' is a Boolean feature, we add another feature ``$ \text{age }< 25 $'' in  $ \mathbf{X} $ by negating the column ``$ \text{age }\ge 25 $''. Hence, in the rest of the paper, we refer $ m $ as the modified number of columns in $ \mathbf{X} $. We next discuss the variables in the MaxSAT problem.

@@

\paragraph{Translating $ E_l, S^i_j, H_l $ to a CNF formula.} 
The soft clauses $ E_l $ and $ S_j^i $ are unit clauses and hence, no translation is required for them.  In the hard clause, when $ y_l=1 $, the simplification is $ H_l:= \neg \eta_l \rightarrow \bigwedge_{i=1}^k {\mathbf{X}_l} \circ {\mathbf{b}_{i}}  $. In this case, we apply the equivalence rule in propositional logic $ (a \rightarrow b) \equiv (\neg a \vee b) $  to encode $ H_l $ into  CNF. In contrast, when $ y_i=0 $, we simplify the hard clause as $ H_l:= \neg \eta_l \rightarrow \neg (\bigwedge_{i=1}^k {\mathbf{X}_l} \circ {\mathbf{b}_{i}}) \Rightarrow  \neg \eta_l \rightarrow \bigvee_{i=1}^k \neg({\mathbf{X}_l} \circ {\mathbf{b}_{i}})$. Since $ \mathbf{X}_l \circ \mathbf{b}_{i} $ constitutes a disjunction of literals (defined in Section~\ref{interpretability_imli_sec:preliminaries}), we apply Tseytin transformation to encode  $ \neg (\mathbf{X}_l \circ \mathbf{b}_{i}) $ into CNF. More specifically, we introduce an auxiliary variable $ z_{l,i} $ corresponding to the clause $ \neg ({\mathbf{X}_l} \circ {\mathbf{b}_{i}}) $. Formally, we replace $ H_l := \neg \eta_l \rightarrow  \bigvee_{i=1}^k \neg ({\mathbf{X}_l} \circ {\mathbf{b}_{i}}) $ with   $ \bigwedge_{i=0}^k H_{l,i} $,  where $ H_{l,0}:= (\neg \eta_l \rightarrow  \bigvee_{i=1}^k  z_{l,i}) $  and  $  H_{l,i}:= z_{l,i} \rightarrow  \neg({\mathbf{X}_l} \circ {\mathbf{b}_{i}})  $ for $ i=\{1,\dots,k\} $. Finally, we  apply the equivalence of $ (a \rightarrow b) \equiv (\neg a \vee b) $ on $ H_{l,i} $ to translate them into CNF. For either value of $ y_l $, the weight of each translated hard clause in the CNF formula is $ \infty $.

@@

\subsection{Learning with Non-binary Features}
	\label{interpretability_imli_sec:non-binary}
	The presented MaxSAT encoding requires input samples to have binary features. Therefore, we discretize datasets with categorical and continuous features into binary features. For each continuous feature, we apply equal-width discretization that splits the feature into a fixed number of bins. For example, consider a continuous feature $ x_c \in [a,b] $. In discretization, we split the domain $ [a,b] $ into three bins with two split points $ \{a',b'\} $ such that $ a<a'<b'<b $.  Therefore, the resulting three discretized features are $ { a \le x_c < a'} $, $ {a' \le x_c < b'} $, and $ { b' \le x_c  \le b } $. An alternate to this close-interval discretization is open-interval discretization, where we consider six discretized features with each feature being compared with one threshold. In that case, the discretized features are: $ x_c \ge a, x_c \ge a', x_c \ge b', x_c < a', x_c < b', x_c \le b $. Both open-interval and close-interval discretization techniques have their use-cases where one or the other is appropriate. For simplicity, we experiment with close-interval discretization in this paper.

	After applying discretization on continuous features, the dataset contains categorical features only, which we convert to binary features using one-hot encoding~\cite{lakkaraju2019faithful,ghosh19imli}. In one-hot encoding, a Boolean vector of features is introduced with cardinality equal to the number of distinct categories. For example, consider a categorical feature having three categories `red', `green', and `yellow'. In one hot encoding, samples with category-value `red', `green', and `yellow' would be converted into binary features by taking values $ 100 $, $ 010 $, and $ 001 $, respectively.

>>>>

\subsection{Description of Variables}
\label{interpretability_imli_sec:variables} 
We initially preprocess $ \mathbf{X} $  to account for the negation of  Boolean features while learning a classifier.\footnote{This preprocessing is similarly applied in  \cite{malioutov2013exact}.} In the preprocessing step, we negate each column in $ \mathbf{X} $ to a new column and append it to $ \mathbf{X} $. For example, if ``$ \text{age }\ge 25 $'' is a Boolean feature, we add another feature ``$ \text{age }< 25 $'' in  $ \mathbf{X} $ by negating the column ``$ \text{age }\ge 25 $''. Hence, in the rest of the chapter, we refer $ m $ as the modified number of columns in $ \mathbf{X} $. We next discuss the variables in the MaxSAT problem.

@@

\paragraph{Translating $ E_l, S^i_j, H_l $ to a CNF formula.} 
The soft clauses $ E_l $ and $ S_j^i $ are unit clauses and hence, no translation is required for them.  In the hard clause, when $ y_l=1 $, the simplification is $ H_l:= \neg \eta_l \rightarrow \bigwedge_{i=1}^k {\mathbf{X}_l} \circ {\mathbf{b}_{i}}  $. In this case, we apply the equivalence rule in propositional logic $ (a \rightarrow b) \equiv (\neg a \vee b) $  to encode $ H_l $ into  CNF. In contrast, when $ y_i=0 $, we simplify the hard clause as $ H_l:= \neg \eta_l \rightarrow \neg (\bigwedge_{i=1}^k {\mathbf{X}_l} \circ {\mathbf{b}_{i}}) \Rightarrow  \neg \eta_l \rightarrow \bigvee_{i=1}^k \neg({\mathbf{X}_l} \circ {\mathbf{b}_{i}})$. Since $ \mathbf{X}_l \circ \mathbf{b}_{i} $ constitutes a disjunction of literals, we apply Tseytin transformation to encode  $ \neg (\mathbf{X}_l \circ \mathbf{b}_{i}) $ into CNF. More specifically, we introduce an auxiliary variable $ z_{l,i} $ corresponding to the clause $ \neg ({\mathbf{X}_l} \circ {\mathbf{b}_{i}}) $. Formally, we replace $ H_l := \neg \eta_l \rightarrow  \bigvee_{i=1}^k \neg ({\mathbf{X}_l} \circ {\mathbf{b}_{i}}) $ with   $ \bigwedge_{i=0}^k H_{l,i} $,  where $ H_{l,0}:= (\neg \eta_l \rightarrow  \bigvee_{i=1}^k  z_{l,i}) $  and  $  H_{l,i}:= z_{l,i} \rightarrow  \neg({\mathbf{X}_l} \circ {\mathbf{b}_{i}})  $ for $ i=\{1,\dots,k\} $. Finally, we  apply the equivalence of $ (a \rightarrow b) \equiv (\neg a \vee b) $ on $ H_{l,i} $ to translate them into CNF. For either value of $ y_l $, the weight of each translated hard clause in the CNF formula is $ \infty $.

@@

\subsection{Learning with Non-binary Features}
	\label{interpretability_imli_sec:non-binary}
	The presented MaxSAT encoding requires input samples to have binary features. Therefore, we discretize datasets with categorical and continuous features into binary features. For each continuous feature, we apply equal-width discretization that splits the feature into a fixed number of bins. For example, consider a continuous feature $ x_c \in [a,b] $. In discretization, we split the domain $ [a,b] $ into three bins with two split points $ \{a',b'\} $ such that $ a<a'<b'<b $.  Therefore, the resulting three discretized features are $ { a \le x_c < a'} $, $ {a' \le x_c < b'} $, and $ { b' \le x_c  \le b } $. An alternate to this close-interval discretization is open-interval discretization, where we consider six discretized features with each feature being compared with one threshold. In that case, the discretized features are: $ x_c \ge a, x_c \ge a', x_c \ge b', x_c < a', x_c < b', x_c \le b $. Both open-interval and close-interval discretization techniques have their use-cases where one or the other is appropriate. For simplicity, we experiment with close-interval discretization in this chapter.

	After applying discretization on continuous features, the dataset contains categorical features only, which we convert to binary features using one-hot encoding~\cite{lakkaraju2019faithful}. In one-hot encoding, a Boolean vector of features is introduced with cardinality equal to the number of distinct categories. For example, consider a categorical feature having three categories `red', `green', and `yellow'. In one hot encoding, samples with category-value `red', `green', and `yellow' would be converted into binary features by taking values $ 100 $, $ 010 $, and $ 001 $, respectively.

### 2022-06-02 chapters/interpretability/imli/experiment.tex
minor edits in many parts of the thesis

\paragraph{Summary of Experimental Results.}
To summarize our experimental results, {\imli} achieves the best balance among prediction accuracy, interpretability, and scalability compared to existing interpretable rule-based classifiers. Particularly, compared to the most accurate classifier RIPPER, {\imli} demonstrates on average $ 1\% $ lower prediction accuracy, wherein the accuracy of {\imli} is higher than BRL, CORELS, and BRS in almost all datasets. In contrast, {\imli} generates significantly smaller rules than RIPPER, specifically in large datasets. Moreover, BRL, CORELS, and BRS report comparatively smaller rule-size than {\imli} on average, but with a significant decrease in accuracy.  In terms of scalability, {\imli} achieves the best performance compared to other interpretable and non-interpretable classifiers by classifying datasets with one million samples. While CORELS also scales to such large datasets,  its accuracy is lower than that of {\imli} by at least $ 2\% $ on average.  Therefore, {\imli} is not only scalable but also accurate in practical classification problems while also being interpretable by design.  We additionally analyze the comparative performance of different formulations presented in this paper, where the  incremental approach empirically proves its efficiency than the na\"ive MaxSAT formulation.  Furthermore, we learn and compare the performance of different interpretable representations: decision lists, decision sets, CNF, and DNF formulas using {\imli} and present the efficacy of {\imli} in learning varied interpretable classifiers. Finally, we study the effect of different hyper-parameters in {\imli}, where each hyper-parameter provides a precise control among training time,  prediction accuracy, and rule-sparsity. In the following, we discuss our experimental results in detail.

@@

	\caption[Training time, test error, and rule-size of different formulations in {\imli}]{Comparison of training time, test error, and rule-size among different formulations presented in the paper. In each cactus plot, the incremental formulation {\imli} with both mini-batch and iterative learning demonstrates the best performance in training time and test error than compared two formulations: non-incremental MaxSAT formulation {\imli}-$\mathsf{B}$ and incremental formulation with only mini-batch learning {\imli}-$\mathsf{M}$. In terms of rule-size, {\imli} often generates higher size rules than {\imli}-$\mathsf{M}$.}
	\label{interpretability_imli_fig:different_formulations}
\end{figure}

\paragraph{Comparison among different formulations in {\imli}.}  We compare the performance of different formulations for learning classification rules as presented in this paper. In Figure~\ref{interpretability_imli_fig:different_formulations}, we show cactus plots for assessing training time (in seconds), test error (in percentage), and rule-size among different formulations. In the cactus plot, a point $ (x,y) $ denotes that the formulation yields lower than or equal to $ y $ training time (similarly, test error and rule-size) in $ x $ many instances in each dataset.

>>>>

\paragraph{Summary of Experimental Results.}
To summarize our experimental results, {\imli} achieves the best balance among prediction accuracy, interpretability, and scalability compared to existing interpretable rule-based classifiers. Particularly, compared to the most accurate classifier RIPPER, {\imli} demonstrates on average $ 1\% $ lower prediction accuracy, wherein the accuracy of {\imli} is higher than BRL, CORELS, and BRS in almost all datasets. In contrast, {\imli} generates significantly smaller rules than RIPPER, specifically in large datasets. Moreover, BRL, CORELS, and BRS report comparatively smaller rule-size than {\imli} on average, but with a significant decrease in accuracy.  In terms of scalability, {\imli} achieves the best performance compared to other interpretable and non-interpretable classifiers by classifying datasets with one million samples. While CORELS also scales to such large datasets,  its accuracy is lower than that of {\imli} by at least $ 2\% $ on average.  Therefore, {\imli} is not only scalable but also accurate in practical classification problems while also being interpretable by design.  We additionally analyze the comparative performance of different formulations presented in this chapter, where the  incremental approach empirically proves its efficiency than the na\"ive MaxSAT formulation.  Furthermore, we learn and compare the performance of different interpretable representations: decision lists, decision sets, CNF, and DNF formulas using {\imli} and present the efficacy of {\imli} in learning varied interpretable classifiers. Finally, we study the effect of different hyper-parameters in {\imli}, where each hyper-parameter provides a precise control among training time,  prediction accuracy, and rule-sparsity. In the following, we discuss our experimental results in detail.

@@

	\caption[Training time, test error, and rule-size of different formulations in {\imli}]{Comparison of training time, test error, and rule-size among different formulations presented in the chapter. In each cactus plot, the incremental formulation {\imli} with both mini-batch and iterative learning demonstrates the best performance in training time and test error than compared two formulations: non-incremental MaxSAT formulation {\imli}-$\mathsf{B}$ and incremental formulation with only mini-batch learning {\imli}-$\mathsf{M}$. In terms of rule-size, {\imli} often generates higher size rules than {\imli}-$\mathsf{M}$.}
	\label{interpretability_imli_fig:different_formulations}
\end{figure}

\paragraph{Comparison among different formulations in {\imli}.}  We compare the performance of different formulations for learning classification rules as presented in this chapter. In Figure~\ref{interpretability_imli_fig:different_formulations}, we show cactus plots for assessing training time (in seconds), test error (in percentage), and rule-size among different formulations. In the cactus plot, a point $ (x,y) $ denotes that the formulation yields lower than or equal to $ y $ training time (similarly, test error and rule-size) in $ x $ many instances in each dataset.
